\documentclass[12pt,letter]{article}
\usepackage{geometry}\geometry{top=0.75in}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{xcolor}	% Color words
\usepackage{cancel}	% Crossing parts of equations out
\usepackage{tikz}    	% Drawing 

% Don't indent
\setlength{\parindent}{0pt}
% Function to replace \section with a problem name specifically formatted
\newcommand{\problem}[1]{\vspace{3mm}\Large\textbf{{Problem {#1}\vspace{3mm}}}\normalsize\\}
% Formatting function, like \problem
\newcommand{\ppart}[1]{\vspace{2mm}\large\textbf{\\Part {#1})\vspace{2mm}}\normalsize\\}
% Formatting 
\newcommand{\condition}[1]{\vspace{1mm}\textbf{{#1}:}\normalsize\\}

\begin{document}
\title{CIS 621 Assignment 4}
\author{Steven Walton}
\maketitle
\problem{1}
For the graph $G=(\mathcal{V},\mathcal{E})$, where $\mathcal{V}$ is the set of 
vertices and $\mathcal{E}$ is the set of edges, we define the following
nonlinear integer program, where $w_{i,j}\geq0, \forall(i,j)\in\mathcal{E}$ and 
$k$ is a nonnegative integer:
\begin{align*}
    \sup&\sum\limits_{(i,j)\in\mathcal{E}} w_{i,j}(x_i + x_j - 2x_ix_j)\\
    %\hspace{-3.5cm}s.t. & \sum\limits_{i\in\mathcal{V}}x_i = k\\
    s.t. & \sum\limits_{i\in\mathcal{V}}x_i = k\\
       & x_i\in\{0,1\}, \forall i \in\mathcal{V}
\end{align*}
Show that the following linear program is a relaxation of the above problem:
\begin{align*}
    \sup&\sum\limits_{(i,j)\in\mathcal{E}} w_{i,j}z_{i,j}\\
     s.t.\hspace{1cm} & z_{i,j}\leq x_i + x_j, \forall(i,j)\in\mathcal{E}\\
     & z_{i,j}\leq 2 - x_i - x_j, \forall(i,j) \in \mathcal{E}\\
     & \sum\limits_{i\in\mathcal{V}}x_i = k\\
     & 0 \leq x_i \leq 1, \forall i \in\mathcal{V}\\
     & 0 \leq z_{i,j} \leq 1, \forall(i,j)\in\mathcal{E}
\end{align*}
Also, let $F(x)=\sum\limits_{(i,j)\in\mathcal{E}}w_{i,j}(x_i + x_j - 2x_ix_j)$ be
the objective function of the nonlinear integer program. Show that for any 
$(x,z)$ that is feasible to the linear program, 
$F(x)\geq\frac12\sum\limits_{(i,j)\in\mathcal{E}}w_{i,j}z_{i,j}$
\ppart{1}
If we remember the definition of linear relaxation we see that it is
\[
    x_i\in\{0,1\} \mapsto 0\leq x_i \leq 1
\]
We can see from the given problem that 
\[
    z_{i,j} = x_i + x_j - 2x_ix_j
\]
We may also notice that $z_{i,j} = 0 \forall x_i,x_j\in\{0,1\}$. We can also 
notice that $z_{i,j} = 1-\delta_{ij}$. Clearly this
needs a relaxation so we can approximate a solution. From here we can easily 
check the given linear program is a relaxation of the initial integer program.

It is trivial to see that $z_{i,j}\leq x_i + x_j \because$ the negative term
ensures that $z_{i,j}$ is smaller. Similarly we can see that if we remove the 
$x_ix_j$ term from the 2 that we obtain a maximal value and thus our $z_{i,j}$
has to be smaller than that. Finally, we see that we have the relaxation terms
where $x_{i,j}$ and $z_{i,j}$ are not limited to integer values. Therefore we
can conclude that this is a relaxation of the initial problem.

\ppart{2}
Looking at the objective function, we'll need to solve for the expectation value.
\begin{align*}
    \langle F(x)\rangle &= \langle\sum\limits_{(i,j)\in\mathcal{E}} w_{i,j}z_{i,j}\rangle\\
           &= \sum\limits_{(i,j)\in\mathcal{E}} w_{i,j}\mathbf{Pr}(\mathsf{vertex\ within\ cut}) \\
           &= \sum\limits_{(i,j)\in\mathcal{E}} w_{i,j}\frac12\\
           &= \frac12 \cancelto{1}{\sum\limits_{(i,j)\in\mathcal{E}}w_{i,j}} \\
           &= \frac12
\end{align*}
We can determine that $\mathbf{Pr}(\mathsf{vertx\ within\ cut}) = \frac12$ if
we better define this as the probability that a vertex is on one side of a cut.
This is clearly a bifrication problem and thus the value is $\frac12$.

Since we can see that the expectation value, $\langle F(x)\rangle = \frac12$, 
and thus we can conclude that $F(x)\geq \frac12\sum\limits_{(i,j)\in\mathcal{E}}w_{i,j}z_{i,j}$

\problem{2}
For the directed graph $G=(\mathcal{V,E})$, where $\mathcal{V}$ is the set of 
vertices and $\mathcal{E}$ is the set of directed edges, we want to partition 
$\mathcal{V}$ into two sets, $\mathcal{U}$ and $\mathcal{W}=\mathcal{V}/
\mathcal{U}$, in order to maximize the total weight of the edges going from 
$\mathcal{U}$ to $\mathcal{W}$ (edges $(i,j)$ with $i\in\mathcal{U}$ and 
$j\in\mathcal{W}$

\begin{itemize}
    \item Give a randomized $\frac14$-approximation algorithm for this problem.
    \item Show that the following linear program is a relaxation of this problem.
\end{itemize}
\begin{align*}
    max & \sum\limits_{(i,j)\in\mathcal{E}} w_{i,j}z_{i,j}\\
    s.t.\hspace{1em}& z_{i,j} \leq x_i, \forall(i,j)\in \mathcal{E}\\
        & z_{i,j}\leq 1 - x_j, \forall(i,j)\in \mathcal{E}\\
        & 0 \leq x_i \leq 1, \forall i\in \mathcal{V}\\
        & 0 \leq z_{i,j} \leq 1, \forall (i,j)\in\mathcal{E}
\end{align*}
\begin{itemize}
    \item For the above linear program, give a randomized $\frac12$-approximation
          algorithm based on rounding $x_i\forall i\in\mathcal{V}$ to 1, with the
          probability of $\frac12x_i + \frac14$
\end{itemize}

\ppart{1}

\ppart{2}

\ppart{3}



\end{document}
